### 2020-10-06 paper.Rmd
minor

For a data matrix $\mathbf{X}_{N \times P} = \{x_{ij}\}_{N \times P}$, 
k-means clustering aims to minimize the error sum of squares (SSE)---the 
within-group variance---which is the sum of the squared Euclidean 
distances between each data point and its cluster centroid. Conversely, 
k-means anticlustering aims to maximize the same objective. The cluster 
centroid is given by averaging the values on each variable over the 
elements within that cluster. The centroid of the *j*th variable in the 
*k*th cluster is given as:

@@

The k-means objective function is then given as

@@

If the $SSE$ is maximal---which is the goal of k-means 
anticlustering---the cluster centroids are as close as possible to the 
overall centroid $\overline{\mathbf{x}}$ and therefore to each other 
[@spath1986]. Thus, k-means anticlustering directly optimizes the 
similarity of the mean values of the $P$ attributes across the $K$ 
clusters.

@@

**This is old, adjust to Steinley (2006):**

@@

# Plot relationship between objectives
layout(matrix(c(1,2,3,3), 2, 2, byrow = TRUE), heights = c(1, 1.4))
plot(
  df[, c("diff_mean", "kmeans")],
  xlab = "mean(f1) - mean(f2)",
  ylab = "SSE",
  pch = 4, 
  cex = .5,
  col = "darkgrey"
)
legend("topright", legend = "A", cex = 1.3, bty = "n")
plot(
  df[, c("diff_var", "kvar")],
  xlab = "var(f1) - var(f2)",
  ylab = "SSE_var",
  pch = 4, 
  cex = .5,
  col = "darkgrey"
)
legend("topright", legend = "B", cex = 1.3, bty = "n")

>>>>

For a data matrix $\mathbf{X}_{N \times P} = \{x_{ij}\}_{N \times P}$, k-means 
clustering aims to minimize the error sum of squares (SSE), also known as the 
within-group variance. The SSE is the sum of the squared Euclidean distances 
between each data point and its cluster centroid. The cluster centroid is given 
by averaging the values on each variable over the elements within that cluster, 
where the centroid $\overline{x}_j ^{(k)}$ of the *j*th variable in the *k*th cluster is

@@

The SSE is then given as:

@@

If the $SSE$ is maximal, the cluster centroids are as close as possible to the 
overall centroid $\overline{\mathbf{x}}$ and therefore to each other 
[@spath1986]. Thus, k-means anticlustering directly optimizes the 
similarity of the mean values of the attributes across the different clusters.

@@

K-means anticlustering, i.e., maximizing SSE minimizes differences between the 
means of the input variables. I present a rather simple extension of the SSE 
criterion to simultaneously minimize differences with regard to the variances 
of the input variables. The variance is defined as the sum of the squared 
distances between all

@@

# Plot relationship between objectives
layout(matrix(c(1,2,3,3), 2, 2, byrow = TRUE), heights = c(1, 1.4))
plot(
  df[, c("diff_mean", "kmeans")],
  xlab = "mean(f1) - mean(f2)",
  ylab = "SSE",
  pch = 4, 
  cex = .5,
  col = "darkgrey"
)
legend("bottomleft", legend = "A", cex = 1.3, bty = "n")
plot(
  df[, c("diff_var", "kvar")],
  xlab = "var(f1) - var(f2)",
  ylab = "SSE_var",
  pch = 4, 
  cex = .5,
  col = "darkgrey"
)
legend("bottomleft", legend = "B", cex = 1.3, bty = "n")
